---
title: "exercise-11"
author: "John Hinkle"
format: html
self-contained: true
editor: visual
---

## Load packages and data

```{r}
library(tidyverse)
library(skimr)
d <- read_tsv("https://raw.githubusercontent.com/difiore/ada-datasets/main/Mammal_lifehistories_v2.txt")
head(d)
```

## Exploratory analysis

According to skim(), there are 10 numeric variables- but I assume that there are actually 9 as "refs"(reference number?) is not intended to be a numeric variable.

```{r}
skim(d)
```

## Step 1: Replace -999 with NA

```{r}
d <- naniar::replace_with_na_all(data = d, condition = ~.x == -999)
```

## Step 2: Drop "litter size" and "refs" variables

```{r}
d <- d |>
  select(c(-`litter size`, -refs))
```

## Step 3: Log transform numeric variables

```{r}
d <- d |>
  mutate(across(where(is.numeric), log))
```

## Step 4a: Run linear models to obtain residuals

```{r}
m1 <- lm(data = d, `gestation(mo)` ~ `mass(g)`, na.action=na.exclude)
m2 <- lm(data = d, `weaning(mo)` ~ `mass(g)`, na.action=na.exclude)
m3 <- lm(data = d, `AFR(mo)` ~ `mass(g)`, na.action=na.exclude)
m4 <- lm(data = d, `AFR(mo)` ~ `mass(g)`, na.action=na.exclude)
m5 <- lm(data = d, `newborn(g)` ~ `mass(g)`, na.action=na.exclude)
m6 <- lm(data = d, `wean mass(g)` ~ `mass(g)`, na.action=na.exclude)
```

## Step 4b: Add new variables to the data set

```{r}
d <- d |>
  mutate(relGest = residuals(m1),
         relWean = residuals(m2),
         relAFR = residuals(m3),
         relLife = residuals(m4),
         relNewbornMass = residuals(m5),
         relWeaningMass = residuals(m6))
```

## Step 5a: Relative maximum lifespan by Order

Primates have the highest relative maximum lifespan.

```{r}
ggplot(data = d, aes(x = order, y = relLife)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Step 5b: Relative newborn mass by Order

Cetacea has the highest relative newborn mass.

```{r}
ggplot(data = d, aes(x = order, y = relNewbornMass)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Step 5c: Relative weaning mass by Order

Perissodactyla has the highest relative weaning mass.

```{r}
ggplot(data = d, aes(x = order, y = relWeaningMass)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Step 6: Run linear models

### Maximum Lifespan

The best model to predict maximum lifespan is **"max. life(mo) \~ gestation(mo) + mass(g) + wean mass(g) + weaning(mo) + 1"**

3 models have a Delta AICc of \<4, plus the "best model"

Gestation period and weaning age are present in all four "top" models.

```{r}
d <- drop_na(d)

library(MuMIn)
global_m1 <- lm(data = d, `max. life(mo)` ~ 
          `gestation(mo)` + 
          `newborn(g)` + 
          `weaning(mo)` +
          `wean mass(g)` + 
          `mass(g)`, 
          na.action = "na.fail")

models1 <- dredge(global_m1, m.lim = c(0, 4))
(models1.list <- get.models(models1, subset = delta < 4))
```

### Plot model-averaged coefficients for the top models

```{r}
models1.avg <- summary(model.avg(models1, subset = delta < 4, fit = TRUE))
plot(models1.avg, full = TRUE, intercept = FALSE)
```

### Age at first reproduction

The best model is **"AFR(mo) \~ gestation(mo) + mass(g) + weaning(mo) + 1"**

6 models have a Delta AICc of \<4, plus the "best model".

Gestation period and weaning age are present in all seven "top" models.

```{r}
global_m2 <- lm(data = d, `AFR(mo)` ~ 
          `gestation(mo)` + 
          `newborn(g)` + 
          `weaning(mo)` +
          `wean mass(g)` + 
          `mass(g)`, 
          na.action = "na.fail")

models2 <- dredge(global_m2, m.lim = c(0, 4))
(models2.list <- get.models(models2, subset = delta < 4))
```

### Plot model-averaged coefficients for the top models

```{r}
models2.avg <- summary(model.avg(models2, subset = delta < 4, fit = TRUE))
plot(models2.avg, full = TRUE, intercept = FALSE)
```
